---
title: "EDA Zindi competition"
author: "Kamau Kamau"
date: "10/05/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,message = F,warning  = F)
```

```{r}
library(tidyverse)
library(knitr)
df = read.csv("Data//Train.csv")
head(df) %>% kable()
```

# The start and end dates of the policy

Most policies in the data set are roughly one year (364 days) with a few exceptions.

```{r}
library(lubridate)
df$Policy.Start.Date <- as.Date(df$Policy.Start.Date)
df$Policy.End.Date   <- as.Date(df$Policy.End.Date)
df$policy_length     <- as.numeric(df$Policy.End.Date - df$Policy.Start.Date)
df <- df %>% filter(policy_length<4000)
df %>% mutate(len = ifelse(policy_length<364,"Less_1_yr",
                           ifelse(policy_length<367,"Appx 1 yr","1yr+")))%>% group_by(len) %>% tally() %>% ggplot(aes(x = len,y = n))+geom_bar(stat = "identity",fill = "green")+coord_flip()+labs(title = "Approximate Policy Duration")
```

# The Gender

The main gender is male at 63% followed be females at 27.55 percent and the others accounting for roughly 9%. The data was cleaned such that gender was "Male","Female" and "Others".

```{r}
library(janitor)
x <- df$Gender %>% tabyl() %>% mutate(perc = paste(round(percent*100,2),"%"))
names(x) <- c("index","n","percent","perc")
x$label = paste(x$index," = ", x$perc)

pie(x$n,labels = x$label,clockwise = F)

df <- df %>% mutate(Gender = ifelse(Gender=="Male","Male",
                                    ifelse(Gender=="Female","Female","Other")))
x <- df$Gender %>% tabyl() %>% mutate(perc = paste(round(percent*100,2),"%"))
names(x) <- c("index","n","percent","perc")
x$label = paste(x$index," = ", x$perc)

pie(x$n,labels = x$label,clockwise = F)

```
# Age

There was several negative age, these was removed.Also any Ages that were less than 10 were removed since it doesnt appear right to have someone claiming insurance under that age. There were also ages that were larger than 100, these also didnt seem right, this remover around 1000 data points, and this decision can be reviewed later. The mean age is between 30 and 50 years

```{r}
df <- df %>% filter(Age>10)
df <- df %>% filter(Age<121)
max(df$Age)
hist(df$Age)
```

# first transaction data & No_Pol

this was converted to a date field. Th majority of respondents hav just one policy(over 70%) and like 15% with 2 policies with a few isolated cases with more policies.

```{r}
df$First.Transaction.Date <- as.Date(df$First.Transaction.Date)
hist(df$No_Pol)
```

# Car Category

Saloons are the most frequent vehicle category at 50.98% followed by vehicles without any stated category and jeeps coming third. We hay have to impute the blanks during further data cleaning because these are very many observations to drop (3340 rows). 

```{r}
x <- df$Car_Category %>% tabyl() %>% mutate(perc = paste(round(percent*100,2),"%"))
names(x) <- c("index","n","percent","perc")
x$label = paste(x$index," = ", x$perc)

pie(x$n,labels = x$label,clockwise = F)
df %>% filter(Car_Category == "")
```
# Vehicle color

The blanks in vehicle color were too many (over 57.45), this column needs to be dropped.

```{r}
x <- df$Subject_Car_Colour %>% tabyl() %>% mutate(perc = paste(round(percent*100,2),"%"))
names(x) <- c("index","n","percent","perc")
x$label = paste(x$index," = ", x$perc)

pie(x$n,labels = x$label,clockwise = F)
```

# Make of the car

These may have to be imputed as well since they had alot of blanks (2307). The main car type/make was toyota at 41.6%.

```{r}
x <- df$Subject_Car_Make %>% tabyl() %>% mutate(perc = paste(round(percent*100,2),"%"))
names(x) <- c("index","n","percent","perc")
x$label = paste(x$index," = ", x$perc)

pie(x$n,labels = x$label,clockwise = F)
df %>% filter(Subject_Car_Make == "")
```

# LGA_Name and State

LGA name is where the policy was purchased while state is the state where the policy was purchased, these have alot of blanks, over 50% and can be removed. For the first model we shall have only the rows that have this variable added. Lagos was the main state where the policies were purchased from.

```{r}
x <- df$LGA_Name %>% tabyl() %>% mutate(perc = paste(round(percent*100,2),"%"))
names(x) <- c("index","n","percent","perc")
x$label = paste(x$index," = ", x$perc)

pie(x$n,labels = x$label,clockwise = F)

x <- df$State %>% tabyl() %>% mutate(perc = paste(round(percent*100,2),"%"))
names(x) <- c("index","n","percent","perc")
x$label = paste(x$index," = ", x$perc)

pie(x$n,labels = x$label,clockwise = F)

```

# Product Name

The main product name was the Car classic at 55% followed by car safe at 31%.

```{r}
x <- df$ProductName %>% tabyl() %>% mutate(perc = paste(round(percent*100,2),"%"))
names(x) <- c("index","n","percent","perc")
x$label = paste(x$index," = ", x$perc)

pie(x$n,labels = x$label,clockwise = F)
```

# Target

The majority of policy holders did not claim insurance, these were as can be expected at 88.19% while those that claimed were 11.81%.

```{r}
x <- df$target %>% tabyl() %>% mutate(perc = paste(round(percent*100,2),"%"))
names(x) <- c("index","n","percent","perc")
x$label = paste(x$index," = ", x$perc)

pie(x$n,labels = x$label,clockwise = F)
```

# The Modelling strategy

* We shall have two datasets, one with all the columns that had few blanks, we shall make a model using these and see how the model performs.

* We shall drop the rows with blanks and train a model that uses the full variables to see the performance

* We shall create an imputed dataset to see how this performs as well

For each of these strategies we shall try to balance the data using over and under sampling which means we shall have a total of 6 cleaned datasets.

```{r}

```

